---
title: "P8131 HW2"
output: pdf_document
author: Shihui Zhu, sz3029
---

```{r library, include=FALSE}
library(ggplot2)
library(tidyverse)
```

## Problem 1

### a) Fill out the table and give comments

Fit the model with logit, probit, and complementary log-log links

```{r input, echo=FALSE}
data1 = tibble(dose_x = 0:4, num_of_dying = c(2, 8, 15, 23, 27), 
               total_num = 30,
               live = total_num-num_of_dying)
data1
```

Fit $g(P(dying)) = \alpha+\beta X$ using logit:

```{r logit}
# Fit GLM
fit.logit = glm(cbind(num_of_dying, live) ~ dose_x, family = binomial(link = 'logit'),
          data = data1)
fit.logit$coefficients
```

So the fitted logit model is 

$$\hat{\pi}(x) = \frac{e^{-2.323790+1.161895x}}{1-e^{-2.323790+1.161895x}}$$
$\beta$ is 1.161895.

Fit $g(P(dying)) = \alpha+\beta X$ using probit:
```{r probit}
fit.probit = glm(cbind(num_of_dying, live) ~ dose_x, family = binomial(link = 'probit'),
          data = data1)
fit.probit$coefficients
```

So the fitted probit model is:
$$\hat{\pi}(x) = \phi(-1.3770923+0.6863805x)$$
$\beta$ is 0.6863805.

Fit $g(P(dying)) = \alpha+\beta X$ using complementary log-log:
```{r cloglog}
fit.cloglog = glm(cbind(num_of_dying, live) ~ dose_x, family = binomial(link = 'cloglog'),
          data = data1)
fit.cloglog$coefficients
```

So the fitted complementary log-log model is:
$$\hat{\pi}(x) = 1 - e^{-e^{-1.9941520+0.7468193x}}$$
$\beta$ is 0.7468193. 

Table:
```{r cis_probit}
# 95% CI for beta
invfisher.logit <- vcov(fit.logit) # inverse of fisher information matrix
invfisher.probit <- vcov(fit.probit)
invfisher.cloglog <- vcov(fit.cloglog)

# Compute CIs
CI.logit = fit.logit$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
CI.probit = fit.probit$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.probit)))))
CI.cloglog = fit.cloglog$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.cloglog)))))
# Bind columns
out.logit = cbind(CI.logit[-1,,drop = FALSE],
                  sum(residuals(fit.logit,type='deviance')^2), # or fit.logit$deviance
                  predict(fit.logit, newdata = tibble(dose_x = 0.01), type = "response"))
out.probit = cbind(CI.probit[-1,,drop = FALSE],
                   sum(residuals(fit.probit,type = 'deviance')^2),
                   predict(fit.probit, newdata = tibble(dose_x = 0.01), type = "response"))
out.cloglog = cbind(CI.cloglog[-1,,drop = FALSE],
                    sum(residuals(fit.cloglog,type='deviance')^2),
                    predict(fit.cloglog, newdata = tibble(dose_x = 0.01), type = "response"))

# Bind rows
out <- rbind(out.logit, out.probit, out.cloglog)
colnames(out)=c('Estimate of Beta','95% CI lower','95% CI upper', "Deviance", "P(dying|x = 0.01)")
rownames(out)=c('logit', 'probit', 'complementary log-log')

out %>% knitr::kable(digits = 3)
```

### (b) Suppose that the dose level is in natural logarithm scale, estimate LD50 with 90% confidence interval based on the three models.

LD50 Estimate for logit model:
```{r ld50_logit}
# Logit
x_0 <- -coef(summary(fit.logit))[1]/coef(summary(fit.logit))[2]
# Point Estimate
# MASS::dose.p(fit.logit, p = 0.5)
x_0.se <- sqrt(
  t(c(-1/coef(summary(fit.logit))[2],
      coef(summary(fit.logit))[1]/coef(summary(fit.logit))[2]^2)) %*% 
    invfisher.logit %*% 
    c(-1/coef(summary(fit.logit))[2],
      coef(summary(fit.logit))[1]/coef(summary(fit.logit))[2]^2))
# 90% CI
ld50.logit <- c(exp(x_0), exp(x_0 - qnorm(0.95)*x_0.se), exp(x_0 + qnorm(0.95)*x_0.se))
```

LD50 Estimate for probit model:
```{r ld50_probit}
# Probit
x_0 <- -coef(summary(fit.probit))[1]/coef(summary(fit.probit))[2]
# Point Estimate
# MASS::dose.p(fit.probit, p = 0.5)
x_0.se <- sqrt(
  t(c(-1/coef(summary(fit.probit))[2],
      coef(summary(fit.probit))[1]/coef(summary(fit.probit))[2]^2)) %*% 
    invfisher.probit %*% 
    c(-1/coef(summary(fit.probit))[2],
      coef(summary(fit.probit))[1]/coef(summary(fit.probit))[2]^2))
# 90% CI
ld50.probit <- c(exp(x_0), exp(x_0 - qnorm(0.95)*x_0.se), exp(x_0 + qnorm(0.95)*x_0.se))
```

LD50 Estimate for cloglog model:
```{r ld50_cloglog}
# C log-log
x_0 <- (log(-log(0.5))-coef(summary(fit.cloglog))[1])/coef(summary(fit.cloglog))[2]
# Point Estimate
# MASS::dose.p(fit.cloglog, p = 0.5)
x_0.se <- sqrt(
  t(c(-1/coef(summary(fit.cloglog))[2],
      (coef(summary(fit.cloglog))[1] - log(-log(0.5)))/coef(summary(fit.cloglog))[2]^2)) %*%
    invfisher.cloglog %*% 
    c(-1/coef(summary(fit.cloglog))[2],
      (coef(summary(fit.cloglog))[1] - log(-log(0.5)))/coef(summary(fit.cloglog))[2]^2))
# 90% CI
ld50.cloglog <- c(exp(x_0), exp(x_0 - qnorm(0.95)*x_0.se), exp(x_0 + qnorm(0.95)*x_0.se))
```

So the LD50 with 90% confidence interval based on the three models is:
```{r table_90, echo=FALSE}
out_ld50 <- rbind(ld50.logit, ld50.probit, ld50.cloglog)
colnames(out_ld50) = c('Estimate of LD50','90% CI lower','90% CI upper')
rownames(out_ld50)=c('logit', 'probit', 'complementary log-log')

out_ld50 %>% knitr::kable(digits = 3)
```

## Problem 2

### (a) How does the model fit the data?

Fit the model using logit

```{r input_2, echo=FALSE}
data2 <- tibble(amount = seq(10, 90, by = 5), 
                offers = c(4, 6, 10, 12, 39, 36, 22, 14, 10, 12, 8, 9, 3, 1, 5, 2, 1),
                enrolls = c(0, 2, 4, 2, 12, 14, 10, 7, 5, 5, 3, 5, 2, 0, 4, 2, 1),
                rejects = offers - enrolls)
data2
```
```{r logit_2}
# Fit GLM
fit.logit = glm(cbind(enrolls, rejects) ~ amount, family = binomial(link = 'logit'),
          data = data2)
fit.logit$coefficients
```

So the fitted logit model is 

$$\hat{\pi}(x) = \frac{e^{-1.64763837+0.03095043x}}{1-e^{-1.64763837+0.03095043x}}$$
This is grouped data. To evaluate how the model fits the data, we compute pearson chi-squared and deviance for the model. And we test on how the model is close to the full model:

$H_0$: The model is close to the full model

$H_1$: not close to full model

```{r goodness}
G.stat <- sum(residuals(fit.logit, type = 'pearson')^2);G.stat # pearson chisq

dev <- fit.logit$deviance;dev # deviance

# compare with chisq(17-2)
pval = 1 - pchisq(dev,15);pval # fit is not good(over dispersion; lack of covariate)
```

The generalized pearson chi-squared statistics is 8.814299, and the deviance is 10.61271. The p-value is large so we do not reject the null hypothesis. The model fits the data fine. 

### (b) How do you interpret the relationship between the scholarship amount and enrollment rate? What is 95% CI?

```{r interpret}
# No scholarship
exp(coef(summary(fit.logit))[1])
# OR_(n+1/n)
exp(coef(summary(fit.logit))[2])
```

The odds of enrolls when there is zero amount of scholarship is 0.192504. And for a one-unit increase in the amount of scholarship, we expect to see about 3.1% increase in the odds of enrolls among those students who were offered with scholarship. 

The 95% CI is:

```{r q2b}
CI1 = fit.logit$coefficients + kronecker(
  t(c(0,qnorm(0.025),-qnorm(0.025))),
  t(t(sqrt(diag(vcov(fit.logit))))))

out = cbind(exp(CI1)[-1,,drop=FALSE])

colnames(out)=c('Estimate for Beta','95% CI','95% CI')
rownames(out)=c('logit')
out %>% knitr::kable(digits = 3)
```


### (c) How much scholarship should we provide to get 40% yield rate (the percentage of admitted students who enroll?) What is the 95% CI?

```{r}
r_star <- log(0.4/0.6)
invfisher.logit <- vcov(fit.logit)
# Logit
x_0 <- (r_star-coef(summary(fit.logit))[1])/coef(summary(fit.logit))[2]
# Point Estimate
# MASS::dose.p(fit.logit, p = 0.4)
x_0.se <- sqrt(
  t(c(-1/coef(summary(fit.logit))[2],
      (coef(summary(fit.logit))[1] - r_star)/coef(summary(fit.logit))[2]^2)) %*% 
    invfisher.logit %*% 
    c(-1/coef(summary(fit.logit))[2],
      (coef(summary(fit.logit))[1] - r_star)/coef(summary(fit.logit))[2]^2))
# 95% CI
ld50.logit <- cbind(x_0, x_0 - qnorm(0.975)*x_0.se, x_0 + qnorm(0.975)*x_0.se)
colnames(ld50.logit) = c('Estimate for Scholarship Amount for 40% Enrollment Rate',
                         '95% CI Lower',
                         '95% CI Upper')
rownames(ld50.logit) = c('logit')
ld50.logit %>% knitr::kable(digits = 3)
```


